Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle Structure

### 2017-09-21 CCIAQVTS/CCIAQVTS.tex
updates for arxiv

% === All of the Packages I use frequently ===
\usepackage{packages/document_config}

% === Causal structure formatting for tikz ===
\input{figures/causal_structures/formatting}

@@

    \begin{figure}
    \begin{nscenter}
        \begin{minipage}[t]{.48\textwidth}
            \centering
            \scalebox{1.0}{\input{figures/causal_structures/bell_structure}}
            \caption{The Bell structure consisting of two observers $A, B$ together with measurement settings $S_{A}$ and $S_{B}$ respectively. The shared latent variable is labeled $\la$.}
            \label{fig:bell_structure}
        \end{minipage}\hspace{0.04\textwidth}%
        \begin{minipage}[t]{.48\textwidth}
            \centering
            \scalebox{1.0}{\input{figures/causal_structures/triangle_structure}}
            \caption{The Triangle structure consisting of three observable variables $A,B,C$ and three latent variables $X, Y, Z$.}
            \label{fig:triangle_structure}
        \end{minipage}
    \end{nscenter}
    \end{figure}

@@

    \section{Fritz Distribution}
    \label{sec:fritz_distribution}
    \begin{figure}
    \begin{nscenter}
        \scalebox{1.0}{\input{figures/causal_structures/triangle_structure_operational_fritz_bell}}
        \caption{The Triangle structure re-imagined to mimic the Bell structure. The measurement settings $S_{A},S_{B}$ are latent nodes unlike the Bell structure (\cref{fig:bell_structure}).}
        \label{fig:triangle_structure_with_fritz_bell_embedded}
    \end{nscenter}
    \end{figure}

@@

    % \subsection{Parameterizing Quantum Distributions}
    % \begin{figure}
    % \begin{nscenter}
    %         \scalebox{1.0}{\input{figures/causal_structures/triangle_structure_operational}}
    %         \caption{The Triangle structure as modeled by quantum states and measurements.}
    %         \label{fig:triangle_structure_quantum_model}
    % \end{nscenter}
    % \end{figure}
    % An alternative explanation for qualitative similarities between \cref{fig:maximum_violation_I_1} and \cref{fig:fritz_distribution_visualized} is related to the sensitively of the initial parameters to our optimization. Above, it was shown that our parameter space of choice was $2\pi$-periodic in all entries. Somewhat surprisingly, when supplied with randomly sampled initial parameters, all optimization methods consistently converged to saturating $I\tsb{Web}$ instead of violating it. We observed that only parameters close enough to the \textit{Fritz parameters} (i.e. parameters that induce the Fritz distribution) were able to achieve violation of $I\tsb{Web}$.

@@

    \begin{definition}
        \label{def:incidence_matrix}
        The \term{incidence matrix} $M$ for a marginal scenario $\mscenario = \bc{V_1, \dots, V_k}$ is a bit-wise matrix where the columns are indexed by \textit{joint} events $j \in \Events{\jointvar}$ and the rows are events by \textit{marginal} events $m \in \Events{\mscenario}$. The entries of $M$ are populated whenever a marginal event $m$ is a restriction of the joint event $j$.
        \[ M_{m,j} \defined \begin{cases}
            1 & m = j|_{\s{D}\br{m}} \\
            0 & \text{otherwise}
        \end{cases} \]
        The incidence matrix has $\abs{\Events{\jointvar}}$ columns, $\abs{\Events{\mscenario}} = \sum_{i} \abs{\Events{V_i}}$ rows and $k\abs{\Events{\jointvar}}$ non-zero entries.
    \end{definition}
    To illustrate this concretely, consider the following example. Let $\jointvar$ be $3$ binary variables $\jointvar = \bc{A,B,C}$ and $\mscenario$ be the marginal scenario $\mscenario = \bc{\bc{A,B}, \bc{B,C}, \bc{A,C}}$. The incidence matrix for $\mscenario$ is:
    \[ M = \kbordermatrix{
        (A,B,C) \:\: \mapsto & (0,0,0) & (0,0,1) & (0,1,0) & (0,1,1) & (1,0,0) & (1,0,1) & (1,1,0) & (1,1,1) \\
        (A\mapsto0, B\mapsto0) & \kone & \kone & \kzer & \kzer & \kzer & \kzer & \kzer & \kzer \\
        (A\mapsto0, B\mapsto1) & \kzer & \kzer & \kone & \kone & \kzer & \kzer & \kzer & \kzer \\
        (A\mapsto1, B\mapsto0) & \kzer & \kzer & \kzer & \kzer & \kone & \kone & \kzer & \kzer \\
        (A\mapsto1, B\mapsto1) & \kzer & \kzer & \kzer & \kzer & \kzer & \kzer & \kone & \kone \\
        (B\mapsto0, C\mapsto0) & \kone & \kzer & \kzer & \kzer & \kone & \kzer & \kzer & \kzer \\
        (B\mapsto0, C\mapsto1) & \kzer & \kone & \kzer & \kzer & \kzer & \kone & \kzer & \kzer \\
        (B\mapsto1, C\mapsto0) & \kzer & \kzer & \kone & \kzer & \kzer & \kzer & \kone & \kzer \\
        (B\mapsto1, C\mapsto1) & \kzer & \kzer & \kzer & \kone & \kzer & \kzer & \kzer & \kone \\
        (A\mapsto0, C\mapsto0) & \kone & \kzer & \kone & \kzer & \kzer & \kzer & \kzer & \kzer \\
        (A\mapsto0, C\mapsto1) & \kzer & \kone & \kzer & \kone & \kzer & \kzer & \kzer & \kzer \\
        (A\mapsto1, C\mapsto0) & \kzer & \kzer & \kzer & \kzer & \kone & \kzer & \kone & \kzer \\
        (A\mapsto1, C\mapsto1) & \kzer & \kzer & \kzer & \kzer & \kzer & \kone & \kzer & \kone
    } \eq \label{eq:example_incidence}\]
    The incidence matrix acts (to the right) on a vector representing the joint probability $\prob[\jointvar]$ distribution and outputs a vector representing the marginal model $\prob^{\mscenario}$. The \term{joint distribution vector} $\probvec^{\jointvar}$ for a probability distribution $\prob[\jointvar]$ is the vector indexed by joint events $j \in \Events{\jointvar}$ whose entries are populated by the probabilities that $\prob[\jointvar]$ assigns to each joint event: $\probvec^{\jointvar}_j = \prob[\jointvar][j]$. Analogously, \term{marginal distribution vector} $\probvec^{\mscenario}$ for a marginal model $\prob^{\mscenario}$ is the vector whose entries are probabilities over the set of marginal outcomes $\Events{\mscenario}$: $\probvec^{\mscenario}_m = \prob[\Dom{m}][m]$.

@@

    \begin{nscenter}
    \begin{figure}
    \begin{subfigure}[b]{.30\linewidth}
    \scalebox{0.65}{\input{figures/causal_structures/inflated_triangle_structure_spiral}}
    \caption{The Spiral inflation.}\label{fig:spiral_inflation}
    \end{subfigure}
    \begin{subfigure}[b]{.30\linewidth}
    \scalebox{0.65}{\input{figures/causal_structures/inflated_triangle_structure_wagon_wheel}}
    \caption{The Wagon-Wheel inflation.}\label{fig:wagon_wheel_inflation}
    \end{subfigure}
    \begin{subfigure}[b]{.30\linewidth}
    \scalebox{0.65}{\input{figures/causal_structures/inflated_triangle_structure}}
    \caption{The Web Inflation.}\label{fig:the_web_inflation}
    \end{subfigure}
    \caption{Some inflations of the Triangle structure.}
    \label{fig:inflations}
    \end{figure}
    \end{nscenter}

@@

    \begin{nscenter}
    \begin{figure}
    \begin{subfigure}[b]{.23\linewidth}
    \scalebox{0.8}{\input{figures/causal_structures/inflated_triangle_structure_spiral_injectable_highlight_1}}
    \caption{$\AnSub[\graph']{\bc{A_1}}$}\label{fig:spiral_inflation_inj1}
    \end{subfigure}
    \begin{subfigure}[b]{.23\linewidth}
    \scalebox{0.8}{\input{figures/causal_structures/triangle_structure_injectable_highlight_1}}
    \caption{$\AnSub[\graph]{\bc{A}}$}\label{fig:spiral_inflation_iminj1}
    \end{subfigure}
    % \par\bigskip % For vertical spacing https://tex.stackexchange.com/questions/120700/vertical-spacing-using-subcaption-package
    \begin{subfigure}[b]{.23\linewidth}
    \scalebox{0.8}{\input{figures/causal_structures/inflated_triangle_structure_spiral_injectable_highlight_2}}
    \caption{$\AnSub[\graph']{\bc{A_2, C_1}}$}\label{fig:spiral_inflation_inj2}
    \end{subfigure}
    \begin{subfigure}[b]{.23\linewidth}
    \scalebox{0.8}{\input{figures/causal_structures/triangle_structure_injectable_highlight_2}}
    \caption{$\AnSub[\graph]{\bc{A,C}}$}\label{fig:spiral_inflation_iminj2}
    \end{subfigure}
    \caption{Some injectable sets $\Inj[\graph]{\graph'}$ of the Spiral inflation ($\graph'$) and their corresponding images $\ImInj[\graph]{\graph'}$ in the Triangle structure ($\graph$).}
    \label{fig:injectablesets}
    \end{figure}
    \end{nscenter}
    % \begin{definition}
    %     Given a causal structure $\graph$ and an inflation $\graph'$ of $\graph$, define the \term{injectable sets} in $\graph'$ and the \term{images of the injectable sets} in $\graph$, (respectively denoted $\Inj[\graph]{\graph'}$ and $\ImInj[\graph]{\graph'}$), as follows:
    %     \begin{align*}
    %         \Inj[\graph]{\graph'} &\defined \bc{N' \subseteq \nodes' \mid \exists N \subseteq \nodes : \AnSub[\graph']{N'} \sim \AnSub[\graph]{N}} \\
    %         \ImInj[\graph]{\graph'} &\defined \bc{N \subseteq \nodes \mid \exists N' \subseteq \nodes' : \AnSub[\graph']{N'} \sim \AnSub[\graph]{N}}
    %     \end{align*}
    % \end{definition}

@@

    \setlength{\bibsep}{3pt plus 3pt minus 2pt}
    \bibliographystyle{apsrev4-1}
    \nocite{apsrev41Control}
    \bibliography{references}

>>>>

% === Bibliography Style Options ===
\bibliographystyle{apsrev4-1}
\setlength{\bibsep}{3pt plus 3pt minus 2pt}

% === Packages ===
\usepackage{document_config}

@@

    \begin{figure}
    \begin{nscenter}
        \begin{minipage}[t]{.48\textwidth}
            \centering
            \scalebox{1.0}{\includegraphics{figures/causal_structures/bell_structure}}
            \caption{The Bell structure consisting of two observers $A, B$ together with measurement settings $S_{A}$ and $S_{B}$ respectively. The shared latent variable is labeled $\la$.}
            \label{fig:bell_structure}
        \end{minipage}\hspace{0.04\textwidth}%
        \begin{minipage}[t]{.48\textwidth}
            \centering
            \scalebox{1.0}{\includegraphics{figures/causal_structures/triangle_structure}}
            \caption{The Triangle structure consisting of three observable variables $A,B,C$ and three latent variables $X, Y, Z$.}
            \label{fig:triangle_structure}
        \end{minipage}
    \end{nscenter}
    \end{figure}

@@

    \section{Fritz Distribution}
    \label{sec:fritz_distribution}
    \begin{figure}
    \begin{nscenter}
        \scalebox{1.0}{\includegraphics{figures/causal_structures/triangle_structure_operational_fritz_bell}}
        \caption{The Triangle structure re-imagined to mimic the Bell structure. The measurement settings $S_{A},S_{B}$ are latent nodes unlike the Bell structure (\cref{fig:bell_structure}).}
        \label{fig:triangle_structure_with_fritz_bell_embedded}
    \end{nscenter}
    \end{figure}

@@

    % \subsection{Parameterizing Quantum Distributions}
    % \begin{figure}
    % \begin{nscenter}
    %         \scalebox{1.0}{\includegraphics{figures/causal_structures/triangle_structure_operational}}
    %         \caption{The Triangle structure as modeled by quantum states and measurements.}
    %         \label{fig:triangle_structure_quantum_model}
    % \end{nscenter}
    % \end{figure}
    % An alternative explanation for qualitative similarities between \cref{fig:maximum_violation_I_1} and \cref{fig:fritz_distribution_visualized} is related to the sensitively of the initial parameters to our optimization. Above, it was shown that our parameter space of choice was $2\pi$-periodic in all entries. Somewhat surprisingly, when supplied with randomly sampled initial parameters, all optimization methods consistently converged to saturating $I\tsb{Web}$ instead of violating it. We observed that only parameters close enough to the \textit{Fritz parameters} (i.e. parameters that induce the Fritz distribution) were able to achieve violation of $I\tsb{Web}$.

@@

    \begin{definition}
        \label{def:incidence_matrix}
        The \term{incidence matrix} $M$ for a marginal scenario $\mscenario = \bc{V_1, \dots, V_k}$ is a bit-wise matrix where the columns are indexed by \textit{joint} events $j \in \Events{\jointvar}$ and the rows are events by \textit{marginal} events $m \in \Events{\mscenario}$. The entries of $M$ are populated whenever a marginal event $m$ is a restriction of the joint event $j$.
        \[ M_{m,j} \defined \begin{cases}
            1 & m = j|_{\s{D}\br{m}} \\
            0 & \text{otherwise}
        \end{cases} \eq \label{eq:incidence_definition}\]
        The incidence matrix has $\abs{\Events{\jointvar}}$ columns, $\abs{\Events{\mscenario}} = \sum_{i} \abs{\Events{V_i}}$ rows and $k\abs{\Events{\jointvar}}$ non-zero entries.
    \end{definition}
    To illustrate this concretely, consider the following example. Let $\jointvar$ be $3$ binary variables $\jointvar = \bc{A,B,C}$ and $\mscenario$ be the marginal scenario $\mscenario = \bc{\bc{A,B}, \bc{B,C}, \bc{A,C}}$. The incidence matrix for $\mscenario$ is:
    \[ M = \kbordermatrix{
        (A,B,C) \:\: \mapsto & (0,0,0) & (0,0,1) & (0,1,0) & (0,1,1) & (1,0,0) & (1,0,1) & (1,1,0) & (1,1,1) \\
        (A\mapsto0, B\mapsto0) & \kone & \kone & \kzer & \kzer & \kzer & \kzer & \kzer & \kzer \\
        (A\mapsto0, B\mapsto1) & \kzer & \kzer & \kone & \kone & \kzer & \kzer & \kzer & \kzer \\
        (A\mapsto1, B\mapsto0) & \kzer & \kzer & \kzer & \kzer & \kone & \kone & \kzer & \kzer \\
        (A\mapsto1, B\mapsto1) & \kzer & \kzer & \kzer & \kzer & \kzer & \kzer & \kone & \kone \\
        (B\mapsto0, C\mapsto0) & \kone & \kzer & \kzer & \kzer & \kone & \kzer & \kzer & \kzer \\
        (B\mapsto0, C\mapsto1) & \kzer & \kone & \kzer & \kzer & \kzer & \kone & \kzer & \kzer \\
        (B\mapsto1, C\mapsto0) & \kzer & \kzer & \kone & \kzer & \kzer & \kzer & \kone & \kzer \\
        (B\mapsto1, C\mapsto1) & \kzer & \kzer & \kzer & \kone & \kzer & \kzer & \kzer & \kone \\
        (A\mapsto0, C\mapsto0) & \kone & \kzer & \kone & \kzer & \kzer & \kzer & \kzer & \kzer \\
        (A\mapsto0, C\mapsto1) & \kzer & \kone & \kzer & \kone & \kzer & \kzer & \kzer & \kzer \\
        (A\mapsto1, C\mapsto0) & \kzer & \kzer & \kzer & \kzer & \kone & \kzer & \kone & \kzer \\
        (A\mapsto1, C\mapsto1) & \kzer & \kzer & \kzer & \kzer & \kzer & \kone & \kzer & \kone
    } \eq \label{eq:example_incidence}\]
    The incidence matrix acts (to the right) on a vector representing the joint probability $\prob[\jointvar]$ distribution and outputs a vector representing the marginal model $\prob^{\mscenario}$. The \term{joint distribution vector} $\probvec^{\jointvar}$ for a probability distribution $\prob[\jointvar]$ is the vector indexed by joint events $j \in \Events{\jointvar}$ whose entries are populated by the probabilities that $\prob[\jointvar]$ assigns to each joint event: $\probvec^{\jointvar}_j = \prob[\jointvar][j]$. Analogously, \term{marginal distribution vector} $\probvec^{\mscenario}$ for a marginal model $\prob^{\mscenario}$ is the vector whose entries are probabilities over the set of marginal outcomes $\Events{\mscenario}$: $\probvec^{\mscenario}_m = \prob[\Dom{m}][m]$.

@@

    \begin{nscenter}
    \begin{figure}
    \begin{subfigure}[b]{.30\linewidth}
    \scalebox{0.65}{\includegraphics{figures/causal_structures/inflated_triangle_structure_spiral}}
    \caption{The Spiral inflation.}\label{fig:spiral_inflation}
    \end{subfigure}
    \begin{subfigure}[b]{.30\linewidth}
    \scalebox{0.65}{\includegraphics{figures/causal_structures/inflated_triangle_structure_wagon_wheel}}
    \caption{The Wagon-Wheel inflation.}\label{fig:wagon_wheel_inflation}
    \end{subfigure}
    \begin{subfigure}[b]{.30\linewidth}
    \scalebox{0.65}{\includegraphics{figures/causal_structures/inflated_triangle_structure}}
    \caption{The Web Inflation.}\label{fig:the_web_inflation}
    \end{subfigure}
    \caption{Some inflations of the Triangle structure.}
    \label{fig:inflations}
    \end{figure}
    \end{nscenter}

@@

    \begin{nscenter}
    \begin{figure}
    \begin{subfigure}[b]{.23\linewidth}
    \scalebox{0.8}{\includegraphics{figures/causal_structures/inflated_triangle_structure_spiral_injectable_highlight_1}}
    \caption{$\AnSub[\graph']{\bc{A_1}}$}\label{fig:spiral_inflation_inj1}
    \end{subfigure}
    \begin{subfigure}[b]{.23\linewidth}
    \scalebox{0.8}{\includegraphics{figures/causal_structures/triangle_structure_injectable_highlight_1}}
    \caption{$\AnSub[\graph]{\bc{A}}$}\label{fig:spiral_inflation_iminj1}
    \end{subfigure}
    % \par\bigskip % For vertical spacing https://tex.stackexchange.com/questions/120700/vertical-spacing-using-subcaption-package
    \begin{subfigure}[b]{.23\linewidth}
    \scalebox{0.8}{\includegraphics{figures/causal_structures/inflated_triangle_structure_spiral_injectable_highlight_2}}
    \caption{$\AnSub[\graph']{\bc{A_2, C_1}}$}\label{fig:spiral_inflation_inj2}
    \end{subfigure}
    \begin{subfigure}[b]{.23\linewidth}
    \scalebox{0.8}{\includegraphics{figures/causal_structures/triangle_structure_injectable_highlight_2}}
    \caption{$\AnSub[\graph]{\bc{A,C}}$}\label{fig:spiral_inflation_iminj2}
    \end{subfigure}
    \caption{Some injectable sets $\Inj[\graph]{\graph'}$ of the Spiral inflation ($\graph'$) and their corresponding images $\ImInj[\graph]{\graph'}$ in the Triangle structure ($\graph$).}
    \label{fig:injectablesets}
    \end{figure}
    \end{nscenter}
    % \begin{definition}
    %     Given a causal structure $\graph$ and an inflation $\graph'$ of $\graph$, define the \term{injectable sets} in $\graph'$ and the \term{images of the injectable sets} in $\graph$, (respectively denoted $\Inj[\graph]{\graph'}$ and $\ImInj[\graph]{\graph'}$), as follows:
    %     \begin{align*}
    %         \Inj[\graph]{\graph'} &\defined \bc{N' \subseteq \nodes' \mid \exists N \subseteq \nodes : \AnSub[\graph']{N'} \sim \AnSub[\graph]{N}} \\
    %         \ImInj[\graph]{\graph'} &\defined \bc{N \subseteq \nodes \mid \exists N' \subseteq \nodes' : \AnSub[\graph']{N'} \sim \AnSub[\graph]{N}}
    %     \end{align*}
    % \end{definition}

@@

    % \bibliographystyle{apsrev4-1}
    \nocite{apsrev41Control}
    \bibliography{references}
